---
title: "Untitled"
author: "Camilo Abbate"
date: "3/29/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Today's outline

\begin{itemize}

\item Importing data
\begin{itemize}
     \item Excel/CSV
     \item RData
\end{itemize}

\item RProjects


\item Data Types
\begin{itemize}
     \item Characters
     \item Strings
     \item Logicals
     \end{itemize}
		

\item Common Statistics

\item Piping (magrittr package)

\item Some coding to start practicing..
	
\end{itemize}


## Importing Data

`R` works with "libraries" (packages)

\begin{center}
\includegraphics[width=2.72in]{rpacka.jpeg}
\end{center}

## Importing Data: 

Before, just in case:
Clean the enviroment (erasing pre-existing stuff)

```{r, eval=F, echo=T}
rm(list = ls())
```


## Importing Data

```{r}

#example_data <- read_excel("C:\\Users\\...\\filename.xlsx")




```





## R Projects!

Let's create an R Project!

Why?

Extremely convenient for file path management and collaboration!

## Where are we working?

```{r}
getwd()
```

## Imagine we need to load some data from

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

