---
title: "Week 5 Exercises"
author: "Data Hack 2022"
date: "4/28/2022"
output: 
  pdf_document:
header-includes:
  - \usepackage{xcolor}
  - \usepackage{dcolumn}
  - \newcommand{\magenta}[1]{\textcolor{magenta}{#1}}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
  - \fancyhead[LE, LO]{General R stuff}
  - \usepackage{caption}
  - \usepackage{framed}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(usmap)
library(sf)
library(pdftools)
library(tidyverse)
library(mapview)
getwd()
ny_hm <- read_csv('ny_hmfm.csv')
```

# More of PDF Extracting and Spatial Data

In this week, we are going to wrap up the use of `pdftools` and start with spatial data.
You'll need to load the following packages (install them if necessary):

`usmap` \hspace{.5cm} `sf` \hspace{.5cm} `mapview`


## Some exercises to practice:

1. Consider the next map:

```{r}
plot_usmap("counties", fill = "yellow", alpha = 0.25,
           include = c(.south_region))
```

Plot a similar map, where counties are in lightblue and exclude Florida.


2. Use the file `ny_hmfm.csv` and consider the following graph:

```{r}
usmap::plot_usmap(data=ny_hm,include="NY", values = "overall_homeless")
```

Exclude the values of overall_homeless for New York City (counties: Bronx, Kings, New York, Queens and Richmond)
You should get a map that looks like this:

```{r,echo=F,fig.height = 3.4, fig.width = 3.4,fig.align = "center"}
#exclude nyc
usmap::plot_usmap(data=ny_hm[-c(3,24,31,41,43),],include="NY", values = "overall_homeless")
```

Now get rid off the counties: Nassau, Suffolk and Westchester. You should get this map:

```{r,echo=F}
#exclude nyc and Nassau(30), Suffolk(52) and Westchester(60)
usmap::plot_usmap(data=ny_hm[-c(3,24,30,31,41,43,52,60),],include="NY", values = "overall_homeless")

```

Finally, google how to get rid off the legend on the bottom left that says: "map_df[,values]"
(If you prefer, try to locate the legend at the top right of the graph)

```{r,echo=F,fig.height = 3, fig.width = 3,fig.align = "center"}
#exclude nyc and Nassau(30), Suffolk(52) and Westchester(60)
usmap::plot_usmap(data=ny_hm[-c(3,24,30,31,41,43,52,60),],include="NY", values = "overall_homeless") + theme(legend.position='none') 
```

\newpage

### A little bit of work with sf:

3. Explore the `sf dataframe` called `new_york_state` (or `ny_no_geometry`to open it faster). Consider any 2 Coc localities and plot their areas.

```{r,include=F}
#just including some:
load("ny_spatial.RData")
load("coc_NY_state.RData")
coc_NY_state <- coc_NY_state %>% mutate(cocnum = str_sub(coc_name,1,6))
ny_for_maps$cocnum <- tolower(ny_for_maps$cocnum)
#jumping waaay ahead of ourselves

new_york_state <- left_join(ny_for_maps,coc_NY_state,by="cocnum")

```

```{r, echo=F,fig.height = 3, fig.width = 3,fig.align = "center",fig.cap = "Example using NY-522 and NY-523 CoC"}
new_york_state[42:48,"coc_area"] %>% plot
```


4. Do the previous graph but with a new variable that represents federal money per homeless person.


 





