---
title: "Lecture 4: Regex and PDF Scraping"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Set Up

Here are the packages we will be using today:

```{r}
library(tidyverse)
library(pdftools) ## new!
library(tabulizer) ## new!
```

# Regular Expressions

Regular expressions or REGEX are just ways to pattern-match on text. I'll start with a few examples to get us familiarized with REGEX, and then we'll show just how powerful they can be.

First off, regular expressions are used to pattern-match on *characters* or *strings* only! You cannot pattern match on *doubles* or *floats*. Hence, whenever we pattern match, we will want to make sure that the column we are matching on is a character type. 

For demonstration purposes, we'll start out with a vector of words and phrases first, just so we're comfortable. We'll also use the `stringr::str_view_all` function to help us out and see our matches in real-time.

## Matching on letters

```{r}
x <- c("Hello! My name is Michael. My phone number is (805) 914-4285. Michael is my name.",
       "MichaelMichaelTopperTopperTopper")
```

We can match on specific phrases by simply typing out what we see:
```{r}
str_view_all(x, pattern = "Michael")
str_view_all(x, pattern = "michael") ## case sensitive!
```

Notice that we are matching on *every* instance of Michael. Regular expressions are greedy by default (e.g, they will match on as much as they possibly can).

## Matching on numbers

We can match on any number we want by using the `\d` regular expression. However, since `\` is a special escape character in R, we actually need to escape the escape character. Hence, to match on any digit, we will use `\\d`

* \d- matches on any digit (type in `\\d` for it to work)

```{r}
str_view_all(x, "\\d")
```
Now let's talk about quantifiers. Suppose we only wanted to match on exactly 3 numbers in a row:

* {n} - match on exactly n number of the previous expression
* {n,m} - match on n to m number of previous expression
* + - match on 1 or more of the previous expression

```{r}
str_view_all(x, "\\d{3}")
str_view_all(x, "\\d{4}") ## only matches on 4 numbers in a row
str_view_all(x, "\\d+")
str_view_all(x, "Michael+")
```

We can use:

* `[a-z]` to match on any letter, or `[A-Z]` to match on any uppercase letter:

```{r}
str_view_all(x, "[a-z]")
str_view_all(x, "[A-Z]")
```

We can use:

* `\s` to match on any whitespace (written as `\\s`).

```{r}
str_view_all(x, "\\s")
```

Lastly, we can use:

* `.` to match on any character. Yes. I mean any character.

```{r}
str_view_all(x, ".")
```

## Anchoring

We can get more flexibility by telling the regex to only match at the beginning of a string or the end of a string.

```{r}
x2 <- c("I want candy.", "Who is the GOAT?", "Who is the Who?", "candy ? candy candy")
```
```{r}
str_view_all(x2, "^Who")

str_view_all(x2, "\\?$") 

str_view_all(x2, "candy$")
```

## Special Characters that you must escape

There are a few characters you need to escape:

* If you want to match on a `.` you must do `\\.`
* If you want to match on a `?` you must do `\\?`
* If you want to match on a `(` you must do `\\(`
* If you want to match on a `[` you msut do `\\[`


## Using the OR operator

You can also use the OR operator for regular expressions:

```{r}
str_view_all(x2, "\\.|\\?")
```

## Practice:

Extract the email addresses:

```{r}
practice <- c("phone number is: (805) 9140-4285",
              "phone numberrrrr is (805) 402-4356",
              "their number ((805) 422-4555), is in the bag")
```

```{r}
str_extract_all(practice, "\\(\\d\\d\\d\\)\\s\\d\\d\\d-\\d\\d\\d\\d")
```

```{r}
practice_1 <- c("email: michaeltopper@ucsb.edu",
                "email: miketopper@gmail.com",
                "email: michaeltopper@umail.ucsb.edu")
```

```{r}
str_extract_all(practice_1, "\\s.{1,}.edu$|\\s.{1,}.com$")
```

## In practice:

In practice we're going to be using the `tidyr::extract` and `tidyr::separate` functions. These will perform regex extracting and splitting on a dataframe rather than just small vector.

```{r, eval = F}
horror_movies <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-22/horror_movies.csv")

horror_movies %>% 
  extract(title, into = "year_released", "(\\d{4})") %>% 
  relocate(year_released) %>% 
  extract(movie_run_time, "movie_run_time", "(\\d{1,})")
```

## Going through a PDF:

Now let's focus on scraping a PDF. First off, we're going to use the function: `pdftools::pdf_text`. This function just extracts all the text from a pdf.


```{r, eval = F}
crime_log <- pdf_text(here::here(paste0("lectures/lecture_4/crime_log.pdf")))
```

```{r, eval = F}
crime_log
```

One of the tricks to getting a pdf into the correct format is to break everything by lines. In a pdf, the newline character is `\n`. We are going to use the following functions for splitting/cleaning:

* str_split() - split into a vector by some regex.
* str_trim() - trim whitespace
* unlist() - will make a list into a vector
* which() - gives the index of a TRUE statement
* as_tibble() - will turn a vector into a dataframe (tibble)

## Brief Refresher

Recall that there are ways to index things in R:
```{r, eval = F}
x2[1]
x2[3]
```

There are also lists:

```{r, eval = F}
list1 <- list(x, x2)
list1[1] # notice the difference
list1[[1]] # between both of these
```

Lists are a way to keep dimensionality down. It is important to know how to index lists. There is an important distinction between the double bracket `[[` which gives you the element inside the list, while `[` gives you the element *of* the list. `[` can extract multiple indices, while `[[` can only extract one exact one. 
## Scraping

We'll always start off scraping by splitting the pdf:
```{r, eval = F}
pdf_trimmed <- crime_log %>% 
  str_split("\n") %>% ## splits the pdf by new-lines (\n)
  unlist() %>% ## unlists the list that was creating by splitting
  str_trim() %>% 
  str_to_lower()
```

```{r, eval = F}
incident_index <- pdf_trimmed %>% 
  str_detect("^incident") %>% 
  which()
```

```{r, eval = F}
pdf_trimmed[incident_index] %>% 
  as_tibble() %>% 
  extract(value, into = "incident", regex = ".{1,}:(.{1,})report.{1,}") %>% 
  mutate(incident = str_trim(incident))
```

Continue on...
